%
% now OS is already added in;
% will add the MD algorithm as well as DRK too in the future
% fm will be done in the near future
% 
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Algorithms for Integral Evaluation Over Gaussian Functions}
%
%
%
%
%

In this chapter, we will focus on the electron integrals; which generally has the form as:
\begin{equation}
 V_{ij} = \int \chi_{i}(r)\chi_{j}(r)f(r,r^{'})
\chi_{k}(r^{'})\chi_{l}(r^{'}) dr dr^{'}
\end{equation}
It is already shown in the shell pair data discussion\ref{shell_pair_eq:1}. Here $\chi$ is 
the Gaussian primitive function with arbitrary angular momentum, and $f(r,r^{'})$ is the 
electron operator for one or two electrons. In this chapter, we will discuss the integrals
and its derivatives.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{History for Algorithms of Integral Evaluation}
%
%
%
%
According to the general review given by Peter Gill\cite{gill1994molecular}, the history for
the development of integral algorithm could be divided into three period of times. The first
one is its infant period, at this time only simple SCF calculation could be performed. The 
first algorithm for Gaussian functions was prompted by Boys\cite{SFBoys1950},
then his methodology was subsequently developed by
Shavitt\cite{1963_int_algorithm}, Clementi etc.
\cite{clementi1969study,clementi1972computation}.

In the second generation, the SCF algorithm was mean to be more general. The ``axis-switch''
method of Pople and Hehre(PH)\cite{PH} revolutionized the algorithms of integral in a sense 
that it enables the SCF calculations to be a standard tool for the chemists. However, this method
is constructed primarily for the S and P type of Gaussian functions. Such deficiency motivated
the development of Dupuis-Rys-King(DRK)\cite{DRK1976JCP, DRK1976JCOMP, DRK1983JCOMP} method 
and McMurchie-Davidson(MD)\cite{MD} method to give more general derivation for the integrals.

The third generation of integral algorithm started from the Obara-Saika method(OS)\cite{OS1986,
OS1988} method, it is derived from the idea to solve the derivatives of integral\cite{HB1982,HB1989}. 
Encouraged by its success, a number of new algorithms came out\cite{HGP,gill1989efficient,
gill1990efficient,PRISM,coldprism} which constitutes the fundamental
layer for modern quantum chemistry packages. 

In this chapter, we will mainly focus on the second and third generation of integral algorithms. 
Namely the MD and DRK methods, OS, HGP and PRISM methods.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% this is about the mathematically formulas for integral
\input{./algorithm/technic/integral/direct}
\input{./algorithm/technic/integral/os}
\input{./algorithm/technic/integral/hgp}
\input{./algorithm/technic/integral/fm}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../note"
%%% End:
